---
title: "Harris_Transformation_Script_Combo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
##Loading Libraries 

```{r}
#setwd(************************)

library(readr)
library(dplyr) 
library(ggplot2)
library(caret)
```

## Loading the Dataset 

You can also embed plots, for example:

```{r}

df <- read_csv("Ames_HousePrice.csv")
head(df)
```

## Dealing with Missing Values 

```{r}
df <- df%>% 
  mutate(Alley = ifelse(is.na(Alley) == TRUE, "no_alley", Alley), 
         MasVnrType = ifelse(is.na(MasVnrType) == TRUE, "None", MasVnrType), 
         BsmtQual = ifelse(is.na(BsmtQual) == TRUE, "None", BsmtQual), 
         BsmtCond = ifelse(is.na(BsmtCond) == TRUE, "None", BsmtCond), 
         BsmtExposure = ifelse(is.na(BsmtExposure) == TRUE, "None", BsmtExposure), 
         BsmtFinType1 = ifelse(is.na(BsmtFinType1) == TRUE, "None", BsmtFinType1), 
         BsmtFinType2 = ifelse(is.na(BsmtFinType2) == TRUE, "None", BsmtFinType2), 
         LotFrontage = ifelse(is.na(LotFrontage) == TRUE, 0, LotFrontage), 
         MasVnrArea = ifelse(is.na(MasVnrArea) == TRUE, 0, MasVnrArea), 
         BsmtFinSF1 = ifelse(is.na(BsmtFinSF1) == TRUE, 0, BsmtFinSF1), 
         BsmtFinSF2 = ifelse(is.na(BsmtFinSF2) == TRUE, 0, BsmtFinSF2), 
         BsmtUnfSF = ifelse(is.na(BsmtUnfSF) == TRUE, 0, BsmtUnfSF),
         TotalBsmtSF = ifelse(is.na(TotalBsmtSF) == TRUE, 0, TotalBsmtSF),
         Electrical = ifelse(is.na(Electrical) == TRUE, "SBrkr", Electrical),
         GarageArea = ifelse(is.na(GarageArea) == TRUE, 0, GarageArea),
         BsmtFullBath = ifelse(is.na(BsmtFullBath) == TRUE, 0, BsmtFullBath), 
         BsmtHalfBath = ifelse(is.na(BsmtHalfBath) == TRUE, 0, BsmtHalfBath), 
         FireplaceQu = ifelse(is.na(FireplaceQu) == TRUE, "no_fireplace", FireplaceQu), 
         GarageType = ifelse(is.na(GarageType) == TRUE, "no_garage", GarageType), 
         GarageFinish = ifelse(is.na(GarageFinish) == TRUE, "no_garage", GarageFinish), 
         GarageQual = ifelse(is.na(GarageQual) == TRUE, "no_garage", GarageQual), 
         GarageCond = ifelse(is.na(GarageCond) == TRUE, "no_garage", GarageCond), 
         PoolQC = ifelse(is.na(PoolQC) == TRUE, "no_pool", PoolQC), 
         Fence = ifelse(is.na(Fence) == TRUE, "no_fence", Fence), 
         MiscFeature = ifelse(is.na(MiscFeature) == TRUE, "no_misc", MiscFeature),
         GarageCars = ifelse(is.na(GarageCars) == TRUE, 0, GarageCars)
         )

#imputation of missing garageyrblt values as the same as respective yearbuilt
df$GarageYrBlt <- ifelse(is.na(df$GarageYrBlt), df$YearBuilt, df$GarageYrBlt)

sum(is.na(df))
colSums(is.na(df))
```

#Making the additional Housing Features 

```{r}
df <- df %>% 
  mutate(was_remodeled = ifelse(YearRemodAdd - YearBuilt == 0, 0, 1), 
         stand_scores = abs(SalePrice - mean(SalePrice))/sd(SalePrice), 
         non_lvingArea = LotArea - GrLivArea  
         )

df$was_remodeled <- factor(df$was_remodeled)

Groups <- df %>% 
  group_by(Neighborhood) %>% 
  summarize(med_price = median(SalePrice), 
            cnt = n()) %>% 
  arrange(med_price) %>% 
  mutate(cum_cnt = cumsum(cnt), 
         neigh_group = ntile(cum_cnt, 5))

df <- df %>% 
  left_join(dplyr::select(Groups, Neighborhood, neigh_group), by = 'Neighborhood') 
df$neigh_group <- factor(df$neigh_group)



```

#Converting Variables to Ordinal Factors 

```{r} 
ranks <- c("None", "Po", "Fa", "TA", "Gd", "Ex")
df$BsmtQual <- factor(as.numeric(factor(df$BsmtQual, ranks))) 
df$BsmtCond <- factor(as.numeric(factor(df$BsmtCond, ranks))) 

ranks_ex <- c('None', 'No', 'Mn', 'Av', 'Gd')
df$BsmtExposure <- factor(as.numeric(factor(df$BsmtExposure, ranks_ex)))

ranks_exterior <- c('Po', 'Fa', 'TA', 'Gd', 'Ex')
df$ExterQual <- factor(as.numeric(factor(df$ExterQual, ranks_exterior)))
df$ExterCond <- factor(as.numeric(factor(df$ExterCond, ranks_exterior))) 

ranks_garage <- c("no_garage", "Po", "Fa", "TA", "Gd", "Ex")
df$GarageQual <- factor(as.numeric(factor(df$GarageQual, ranks_garage))) 
df$GarageCond <- factor(as.numeric(factor(df$GarageCond, ranks_garage))) 

ranks_pool <- c("no_pool", "Fa", "TA", "Gd", "Ex")
df$PoolQC <- factor(as.numeric(factor(df$PoolQC, ranks_pool))) 

ranks_3 <- c('Po', 'Fa', 'TA', 'Gd', 'Ex')
df$HeatingQC <- factor(as.numeric(factor(df$HeatingQC, ranks_3)))
df$KitchenQual <- factor(as.numeric(factor(df$KitchenQual, ranks_3))) 

#making orders
ranks_mason <- c('None','BrkCmn', 'BrkFace', 'CBlock', 'Stone') 
df$MasVnrType <- factor(df$MasVnrType, ranks_mason)  

ranks_housestyle <- c('1Story', '1.5Unf', '1.5Fin', '2Story', '2.5Unf', '2.5Fin', 'SFoyer', 'SLvl')  
df$HouseStyle <- factor(df$HouseStyle, ranks_housestyle) 

ranks_bsmttype <- c('None', 'Unf', 'LwQ', 'Rec', 'BLQ', 'ALQ', 'GLQ')
df$BsmtFinType1 <- factor(df$BsmtFinType1, ranks_bsmttype) 
df$BsmtFinType2 <- factor(df$BsmtFinType2, ranks_bsmttype)

ranks_fence <- c("no_fence", "MnWw", "GdWo", "MnPrv", "GdPrv")
df$Fence <- factor(df$Fence, ranks_fence) 

ranks_fireplace <- c("no_fireplace", "Po", "Fa", "TA", "Gd", "Ex")
df$FireplaceQu <- factor(df$FireplaceQu, ranks_fireplace)

ranks_functional <- c("Sal", "Sev", "Maj2", "Maj1", "Mod", "Min2", "Min1", "Typ")
df$Functional <- factor(df$Functional, ranks_functional)

ranks_garagefinish <- c("no_garage", "Unf", "RFn", "Fin")
df$GarageFinish <- factor(df$GarageFinish, ranks_garagefinish)

ranks_drive <- c("N", "P", "Y")
df$PavedDrive <- factor(df$PavedDrive, ranks_drive)

ranks_saletype <- c("Oth", "ConLD", "ConLI", "ConLw", "Con", "COD", "New", "VWD", "CWD", "WD")
df$SaleType <- factor(df$SaleType, ranks_saletype)

ranks_salecondition <- c("Partial", "Family", "Alloca", "AdjLand", "Abnorml", "Normal")
df$SaleCondition <- factor(df$SaleCondition, ranks_salecondition)

```

#Making factored dummy variables for all other categorical variables

```{r}

df$MSZoning <- factor(df$MSZoning)
df$Street <- factor(df$Street)
df$Alley <- factor(df$Alley) 
df$LotShape <- factor(df$LotShape) 
df$LandContour <- factor(df$LandContour) 
df$Utilities <- factor(df$Utilities) 
df$LotConfig <- factor(df$LotConfig) 
df$LandSlope <- factor(df$LandSlope) 
df$Condition1 <- factor(df$Condition1) 
df$Condition2 <- factor(df$Condition2) 
df$BldgType <- factor(df$BldgType) 
df$RoofStyle <- factor(df$RoofStyle) 
df$RoofMatl <- factor(df$RoofMatl) 
df$Exterior1st <- factor(df$Exterior1st) 
df$Exterior2nd <- factor(df$Exterior2nd) 
df$MasVnrType <- factor(df$MasVnrType) 
df$Foundation <- factor(df$Foundation) 
df$HouseStyle <- factor(df$HouseStyle)  
df$Neighborhood <- factor(df$Neighborhood) 
df$Heating <- factor(df$Heating)
df$Electrical <- factor(df$Electrical) 
df$GarageType <- factor(df$GarageType) 
df$PavedDrive <- factor(df$PavedDrive) 
df$MiscFeature <- factor(df$MiscFeature) 
df$SaleType <- factor(df$SaleType) 
df$SaleCondition <- factor(df$SaleCondition) 
df$CentralAir <- factor(df$CentralAir) 
```

#removing outliers

```{r} 
df1 <- df %>% 
  filter(stand_scores < 3 & LotArea < 25000)

names(df1)[names(df1) == "1stFlrSF"] <- "X1stFlrSF"
names(df1)[names(df1) == "2ndFlrSF"] <- "X2ndFlrSF"
names(df1)[names(df1) == "3SsnPorch"] <- "X3SsnPorch"

df2 <- df1 %>% 
  dplyr::select(-PID, -X1, -GrLivArea, -stand_scores, -SaleType, -SaleCondition, -MoSold, -YrSold, 
                -GarageCars, -GarageYrBlt, -TotRmsAbvGrd, -TotalBsmtSF, -YearBuilt, -YearRemodAdd,  
                -OverallCond, -OverallQual, -Condition1, -Condition2, -LotFrontage, -LotArea)



```
#scaling the area and square footage
```{r}

df3 <- df2 %>% 
  mutate(SalePrice = log(SalePrice + 1),
         MasVnrArea = log(MasVnrArea + 1), 
         BsmtFinSF1 = log(BsmtFinSF1 + 1), 
         BsmtFinSF2 = log(BsmtFinSF2 + 1), 
         BsmtUnfSF = log(BsmtUnfSF + 1),
         X1stFlrSF = log(X1stFlrSF + 1), 
         X2ndFlrSF = log(X2ndFlrSF + 1), 
         GarageArea = log(GarageArea + 1),
         WoodDeckSF = log(WoodDeckSF + 1),
         OpenPorchSF = log(OpenPorchSF + 1),
         EnclosedPorch = log(EnclosedPorch + 1),
         X3SsnPorch = log(X3SsnPorch + 1),
         ScreenPorch = log(ScreenPorch + 1),
         PoolArea = log(PoolArea + 1),
         non_lvingArea = log(non_lvingArea + 1)
         )
#for some reason, an na value was produced for the following entry after scaling
df3$non_lvingArea = df3$non_lvingArea[286] = log(7844 + 1)

```

#variable selection

```{r} 
set.seed(1212)

index = sample(1:nrow(df3), size = nrow(df3)*0.8)

train_data = df3[index, ]
test_data = df3[-index, ]


mdl_null <- lm(SalePrice ~ X1stFlrSF, data = df3)

mdl_full <- lm(SalePrice ~ BldgType + HouseStyle + RoofStyle + 
                   MasVnrType + MasVnrArea + ExterQual + ExterCond + 
                   BsmtQual + BsmtCond + BsmtFinType1 + 
                   BsmtFinSF1 + BsmtUnfSF  + 
                   X1stFlrSF + X2ndFlrSF + 
                   BsmtFullBath + BsmtHalfBath + FullBath + HalfBath + 
                   BedroomAbvGr + KitchenQual  + 
                   GarageType + GarageFinish + GarageArea + 
                   GarageQual + GarageCond + WoodDeckSF + OpenPorchSF + 
                   EnclosedPorch + X3SsnPorch + ScreenPorch + PoolArea + PoolQC + 
                   Fence + MiscFeature + MiscVal + was_remodeled + non_lvingArea, data = df3)

step_BIC1 <- step(mdl_full, k = log(2500)) 

step_BIC2 <- step(mdl_null, scope = list(lower = mdl_null, upper = mdl_full), direction = 'both', k = log(2500))
summary(step_BIC2)

mdl_bic_round_2 <- lm(SalePrice ~ BldgType + 
                   ExterQual + ExterCond + 
                   BsmtQual + BsmtFinType1 + 
                   BsmtFinSF1  + 
                   X1stFlrSF + X2ndFlrSF + 
                   FullBath + HalfBath + 
                   KitchenQual  + 
                   GarageFinish + GarageArea + 
                   GarageCond + WoodDeckSF + Neighborhood +
                   was_remodeled, data = df3)

step_BIC3 <- step(mdl_bic_round_2, k = log(2500)) 

step_BIC4 <- step(mdl_null, scope = list(lower = mdl_null, upper = mdl_bic_round_2), direction = 'both', k = log(2500))
summary(step_BIC4)

mdl_bic_round_3 <- lm(SalePrice ~ BldgType + 
                   ExterQual + ExterCond + 
                   BsmtQual + 
                   BsmtFinSF1  + 
                   X1stFlrSF + X2ndFlrSF + 
                   KitchenQual  + HeatingQC +
                   GarageQual + Fireplace, data = df3)

step_BIC5 <- step(mdl_bic_round_3, k = log(2500)) 

step_BIC6 <- step(mdl_null, scope = list(lower = mdl_null, upper = mdl_bic_round_3), direction = 'both', k = log(2500))
summary(step_BIC6)

mdl_test <- lm(SalePrice ~ BsmtQual + X1stFlrSF + X2ndFlrSF +
                       ExterQual + ExterCond + GarageQual + KitchenQual + 
                       BsmtQual + HeatingQC + Fireplaces + neigh_group, data = df3)

Testing <- step(mdl_null, scope = list(lower = mdl_null, upper = mdl_test), direction = 'both', k = log(2500)) 
summary(Testing)
```
#Cross Validation

```{r}



ctrl = trainControl(method = "cv", number = 10)

lm_modelcv <- train(SalePrice ~ BsmtQual + X1stFlrSF + X2ndFlrSF +
                       ExterQual + ExterCond + GarageQual + KitchenQual + 
                       BsmtQual + HeatingQC + Fireplaces + neigh_group, data=train_data, method = "lm", trControl = ctrl)
```
```{r}
lm_modelcv
```


```